---
title: "Manipulating NACA parameters"
author: "Daniel Kaplan"
date: "7/24/2020"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mosaic)
library(mosaicCalc)
library(ggformula)
```

The NACA 4-digit airfoil framework involves three inputs:

- $x$ which measures the front-back position from 0 up front to 1 at the back.
- $h$ the maximum thickness of the airfoil
- $p$ the front-back position of the highest point of the camber line
- $m$ the height of the highest point of the camber line

For any one airfoil, we fix $h$, $p$, and $m$ at some numerical value and plot out the camber line $y_c(x)$ as a function of $x$. We would also draw the curves of the top and bottom of the airfoil, which are found by adding or subtracting the thickness from the camber line. (The addition is done *perpendicularly* to the camber line, not straight up and down.)

$$ y_c(x) \equiv \left\{
\begin{array}{lc}\frac{m}{p^2}(2 p x - x^2) & 0 \leq x < p\\
\frac{m}{(1-p)^2} (1 - 2 p + 2 p x - x^2)& p \leq x \leq 1\end{array} \right.$$

## A computer implementation of the camber line

We will plot $y_c(x)$ versus $x$ for any selected value for $h$, $p$, and $m$. 

`r sliderInput('p', 'Parameter p', min = 0, max = 1, value = 0.4 )`

`r sliderInput('m', "Parameter m", min = 0, max = 0.2, value = 0.05)`

`r sliderInput('h', "Parameter h", min =0, max = 0.5, value = .2)`

`r plotOutput("camber_plot")`


```{r echo=FALSE}
camber_front <- makeFun(
  m * (2*p*x - x^2)/(p^2) ~ x + m + p)
dcamber <- makeFun(
  ifelse(x < p, 
         2*m*(p-x)/(p^2),
         2*m*(p-x)/((1-p)^2)) ~ x + m + p)
  
camber_back <- makeFun(
  m * (1 - 2*p + 2*p*x - x^2)/((1-p)^2) ~ x + m + p)
thickness <- makeFun(
  5 * h * (0.2969*sqrt(x) - 0.126*x - 0.3516*x^2 + 0.2843*x^3 +
         -0.1036 * x^4) ~ x + h)
camber <- 
  makeFun(ifelse(x < p, 
                 camber_front(x, m=m, p=p), 
                 camber_back(x, m=m, p=p)) ~ x + m + p)
```


```{r echo=FALSE}
output$camber_plot <- renderPlot({
  m <- input$m
  p <- input$p
  h <- input$h
  camber_fun <- makeFun(camber(x, m = m, p = p) ~ x)
  
  x <- (1 - cos(seq(0, pi, length=200)))/ 2
  yc <- camber(x, m = m, p = p)
  dyc <- dcamber(x, m = m, p = p)
  yt <- thickness(x, h = h)
  theta <- atan(dyc)
  Airfoil <- data.frame(
    xpos = x,
    xtop = x + yt * sin(theta),
    ytop = yc + yt * cos(theta),
    xbottom = x - yt * sin(theta),
    ybottom = yc - yt * cos(theta)
  )
  
  slice_plot(camber_fun(x) ~ x, domain(x = c(0, 1)), 
             npts = 300, color = "orange") %>%
    gf_line(ytop ~ xtop, data = Airfoil) %>%
    gf_line(ybottom ~ xbottom, data = Airfoil) %>%
    gf_lims(y = c(-.5, .5))
})
```

## The domain of the function

The airfoil-shape function has a valid output (that is, one that is legitimately an airfoil), only for $0 \leq x \leq 1$ and for some regions of the $(m, p, h)$ parameter space.

Play around with the sliders to find the boundary of the domain in $(m, p, h)$ space. Give a few examples of points on the boundary.

## For later in the course

1. Finding the angle of the thickness by differentiating the camber line with respect to x.
2. Does the partial effect of $p$ depend on $m$? (Yes, since when $m=0$, $p$ has no effect.)
