---
title: "Force-balance dynamics"
author: "Daniel Kaplan"
runtime: shiny
---

```{r include=FALSE}
library(mosaic)
library(ggformula)
library(mosaicCalc)
```

Force-balance dynamics are those reflecting Newton's Second Law of Motion ($F = ma = m\, \partial_{tt} x$). We're going to work with systems where the force itself is a simple function of $x$ and $\partial_t x$:
$$m\, \partial_{tt} x = -b x - a \partial_t x$$ where $a$ and $b$ are constants. The negative sign before the $k$ is a convention that let's positive $a$ and positive $b$ correspond to the physical system of a damped spring, but has no further significance.

Again, motivated by the observed motion of a spring-mass system, we'll check whether solutions of the simple form $x(t) \equiv e^{-kt}\, \sin(\omega t)$ will satisfy the equation. The graph shows the three functions $x(t)$, $\partial_t x(t)$, and $\partial_{tt} x(t)$ for whatever choice you make of $a/m$ and $b/m$.

`r plotOutput("xgraph", height="200px", brush="crosshair")`

`r sliderInput("k", "k", min = -2, max=0, value=-0.5, step = 0.01)`
`r sliderInput("omega", "omega", min = 2, max=2.5, value=1, step = 0.001)`

`r plotOutput("difference", height="200px")`

`r plotOutput("forced", height="200px")`

`r plotOutput("phase", height="200px")`

```{r}
f <- reactive({
  omega <- input$omega
  k <- input$k
  function(t) exp(k*t) * sin(omega*t)
})
df <- reactive({
  mosaicCalc::D(f()(t) ~ t)
})
ddf <- reactive({
  mosaicCalc::D(f()(t) ~ t + t)
})

ff <- reactive({
  omega <- input$omega
  function(t) sin(omega*t)
})
dff <- reactive({
  mosaicCalc::D(ff()(t) ~ t)
})
ddff <- reactive({
  mosaicCalc::D(ff()(t) ~ t + t)
})


output$xgraph <- renderPlot({
  x <- f()
  dx <- df()
  ddx <- ddf()
  ## Figuring out a and b from k and omega
  sum <- function(t) ddx(t) - 2*input$k* dx(t) + 
    ((input$k)^2 + (input$omega)^2) * x(t)
  
  slice_plot(sum(t) ~ t, color="green", size=2, domain(t=c(0,15))) %>%
    slice_plot(dx(t) ~ t, color = "red") %>%
    slice_plot(ddx(t) ~ t, color = "blue") %>%
    slice_plot(x(t) ~ t)
})

output$difference <- renderPlot({
  x <- f()
  dx <- df()
  ddx <- ddf()
  slice_plot(ddx(t) + 5*x(t) + dx(t) ~ t, domain(t=c(0,15)))
})

output$forced <- renderPlot({
  ddx <- ddff()
  dx <- dff()
  x <- ff()
  slice_plot((ddx(t) + 5*x(t) + .02*dx(t)) ~ t, domain(t=c(0,15)))
})

output$phase <- renderPlot({
  slice_plot(180*atan2(omega*0.02, (5 - omega^2))/pi ~ omega, domain(omega=sqrt(5)*c(.95, 1.05)))
})
```

```{r}
amplitude <- function(omega, b, c) {
  1/sqrt((c-omega^2)^2 + omega^2*b^2)
}
slice_plot(amplitude(omega, b=.01, c=5) ~ omega, domain(omega=c(.3, 3)))
```




A solution to the dynamics will be a function $x(t)$ that satisfies the force-balance dynamics. Again, motivated by the motion of a damped spring, 

It's very common for the force on an object to be a function of the object's position $y$ and it's velocity $v = \partial_t y$. For instance, in a spring-mass system, the force exerted by the spring depends on the spring's extension. There's also a force of air resistance or internal friction in the spring material. We'll write the net force as $F(y, v) \equiv F(y, \partial_t y)$.

Like all continuous functions, $F(y, v)$ can be approximated locally by a  first-order polynomial: $$F(y, v) = a +  b y + c v$$ To simplify things even more, let's measure position not directly with $y$ but with $x \equiv y + a/b$. Rewriting the definition for $F()$ in terms of $x$ gives $$F(x, v) = b x + c v$$

Newton's Second Law of motion (F = ma) implies that $$m \partial_{{tt}} x = F(x, v)

x + ystar = y 
