---
title: "Sums of sinusoids"
author: "Daniel Kaplan"
date: "7/31/2020"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
A1  <- sliderInput("A1", "Oscillator 1", 
                   min=0,  max=2, value  = 1, step  = 0.1)
A2  <- sliderInput("A2", "Oscillator 2", min=0,  max=2, value  = 0,
                   step = 0.1)
P1  <- sliderInput("P1", "",  
                   min=0.5, max = 5, value = 2, step=0.1)
P2  <- sliderInput("P2", "",  
                   min=0.5, max = 5, value = 2, step=0.1)
phi1 <- sliderInput("phi1", "", 
                    min = 0, max = 2*pi, value  = 0, step = 0.01)
phi2 <- sliderInput("phi2", "", 
                    min = 0, max = 2*pi, value  = 0, step = 0.01)

```

Color    | Amplitude | Period | Phase
---------|-----------|--------|--------
Blue    | `r A1`    | `r P1` | `r phi1`
Green   | `r A2`    | `r P2` | `r phi2`


`r sliderInput("baseline", "Baseline", min=-2, max = 2, value=0, step=0.01)`

`r  plotOutput("main_plot")`


```{r echo=FALSE}
timebase <- reactive({
  seq(0,20, length = 500)
})

waveforms <- reactive({
  res <- data.frame(
    time = timebase() ,
    first  = input$A1*sin(2*pi*timebase()/input$P1 + input$phi1),
    second = input$A2*sin(2*pi*timebase()/input$P2 + input$phi2)
  ) 
  
  res$sum <-  res$first + res$second + input$baseline
  
  res
})


output$main_plot <- renderPlot({
  ggplot(data = waveforms(), aes(x = time, y = sum)) +
    geom_line(color = "black", size = 3, alpha = 0.25)  +
    geom_line(aes(y = first), color = "blue") +
    geom_line(aes(y  = second), color = "green")  +
    ylim(-4,  4)
})

